---
title: "Project_Red_Wines V2"
author: "LLT"
date: "Wednesday, May 6, 2015"
output: html_document
---
#Analysis
##Preparations
First of all, let's load the data into a dataframe and all necessary libraries

```{r,echo=FALSE,warning=FALSE}
red_wines<-read.csv('wineQualityReds.csv', header = TRUE, sep = ',', 
                    row.names = 1)


library(ggplot2)
library(RColorBrewer)
library(GGally)
```

##First look

```{r}
names(red_wines)
str(red_wines)
summary(red_wines)
```

We have 12 numeric variables. The summary shows max really far out for residual.sugar, chlorides, free.sulfur, density... Those might be outlier/reporting problem or really special wines.

###Quality
First of all, what about quality of those red wines
Also, quality is a numeric, we will add it as a factor.

```{r,echo=FALSE,warning=FALSE}
red_wines$quality.lvl=cut(red_wines$quality, 
                          c(0,1,2,3,4,5,6,7,8,9,10), 
                          labels = c(1,2,3,4,5,6,7,8,9,10))
#now going with a boxplot
ggplot(aes(x = quality.lvl), data = red_wines) +
  geom_bar() +
  scale_x_discrete()
```

It looks like a normal distribution. Notes are out of 10. Most of the wines are around 5/6.
My main line of inquiries will be the relation between all variables to quality.
Looking at alcohol:

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = alcohol), data = red_wines) +
  geom_bar(binwidth = 0.1, position = 'identity')
```

Much more disparate but we have a nice peak around 9.5
Does alcohol level affects  the quality?

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = quality.lvl, y = alcohol), data = red_wines) +
  geom_boxplot(outlier.colour= 'red') +
  ggtitle("Alcohol by quality")

#correlation calculation
cor(red_wines$alcohol, red_wines$quality)
```

There is a medium correlation between quality and alcohol, the graph shows that the higher the quality, the higher the alcohol. This is especially true for the higher end wines.

###Acidity variables
We have 3 types of acidity listed: 

* Fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
* Volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar tast
* Citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines

They should all relate to the ph I think. 

```{r}
#fixed pH and fixed acidity correlation
cor(red_wines$pH, red_wines$fixed.acidity)

#fixed pH and volatile acidity correlation
cor(red_wines$pH, red_wines$volatile.acidity)

#fixed pH and citric acid correlation
cor(red_wines$pH, red_wines$citric.acid)

```

So Fixed acidity and citric acid correlate quite strongly with the pH, but not the volatile acidity.

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = pH, y = volatile.acidity), data = red_wines) +
  geom_point() +
  ggtitle("Volatile acidity by pH")

#modifying alpha for overplotting
ggplot(aes(x = pH, y = volatile.acidity), data = red_wines) +
  geom_jitter(alpha = 0.3) +
  ggtitle("Volatile acidity by pH")
```

Volatile acidity and pH are not related. Even for high quality wines, the pH seems to vary greatly. 

###Quality and pH
Let's get a closer look between those pH and quality now.

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = quality.lvl, y = pH), data = red_wines) +
  geom_boxplot() +
  ggtitle("pH by quality")

```

The higher the quality, the lower the pH it seems.
What about salt?

###Quality and chlorides

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = quality.lvl, y = chlorides), data = red_wines) +
  geom_boxplot(outlier.colour = 'red') +
  ggtitle("Chlorides by quality")
summary(red_wines$chlorides)
```

Outliers are colored in red in the plot. The summary shows that the max is at 0.611 whereas the 3rd quartile is at 0.09. Lets zoom in to have  a better view.

```{r,echo=FALSE,warning=FALSE}
#coloring outlier and limiting the y scale to have a better look

ggplot(aes(x = quality.lvl, y = chlorides), data = red_wines) +
  geom_boxplot(outlier.colour = 'red') +
  coord_cartesian() +
  ylim(quantile(red_wines$chlorides,0.05),quantile(red_wines$chlorides,0.95)) +
  ggtitle("Chlorides by quality, red outliers")
```

The second plot is limited between the 0.05 and the 0.95 quantiles. There is some overlaps, but the higher the quality, the least salt.

```{r,warning=FALSE}
cor(red_wines$quality,red_wines$chlorides)
```

Last variable on the obvious affecting taste (and so quality) I think is sugar

###Quality and residual sugar

```{r,echo=FALSE,warning=FALSE}
#first plot with alpha for overplotting
ggplot(aes(x = quality.lvl, y = residual.sugar), data = red_wines) +
  geom_jitter(alpha = 0.4) +
  scale_y_log10() +
  ggtitle("Residual sugar by quality, log y axe")

#limiting y  axe for lisibility
ggplot(aes(x = quality.lvl, y = residual.sugar), data = red_wines) +
  geom_boxplot(outlier.color='red') +
  coord_cartesian() +
  ylim(quantile(red_wines$residual.sugar,0.05),
       quantile(red_wines$residual.sugar,0.95)) +
  ggtitle("Residual sugar by quality (0.05 to 0.95 quant")
```

No clear relationship between the sugar and quality.

```{r,warning=FALSE}
cor(red_wines$quality,red_wines$residual.sugar)
````

Correlation confirms that there is no relationships.

##Getting a taste
Let's try multi-variables plots to see if we can have a better picture.

```{r,echo=FALSE,warning=FALSE, fig.width=20, fig.height=20}
#ggpairs to see the relationships of all variables by pairs
ggpairs(red_wines)
```

###Chlorides and alcohol by quality

```{r,echo=FALSE,warning=FALSE}
#Fist will be alcohol and chlorides, facetted by quality. 
#A smoother was added to see the overall trend.
ggplot_base_alcohol.quality<-ggplot(aes(x = alcohol), data = red_wines) +
  facet_wrap(~quality.lvl, scales = 'free_y')
  
ggplot_base_alcohol.quality +
  geom_point(aes(y = chlorides)) +
  geom_smooth(aes(y = chlorides)) +
  ggtitle("Chlorides by alcohol facetted by quality")

```

The overall trend is downward: less chlorides as the alcohol level gets higher. 
But we can see that for the top 2 qualities, we got a rebound around 12° where salt level rises a bit. It might be some outlier as we only got a few points for those quality.

```{r,warning=FALSE}
with(subset(red_wines, quality.lvl==5),cor(chlorides,alcohol))
with(subset(red_wines, quality.lvl==6),cor(chlorides,alcohol))
with(subset(red_wines, quality.lvl==8),cor(chlorides,alcohol))

```

Low correlation between alcohol and chlorides.

###Residual sugar and alcohol by quality
Let's look at the same thing with sugar:

```{r,echo=FALSE,warning=FALSE}
ggplot_base_alcohol.quality +
  geom_point(aes(y = residual.sugar)) +
  geom_smooth(aes(y = residual.sugar)) +
  ggtitle("Residual sugar by alcohol facetted by quality")

```

```{r,warning=FALSE}
with(subset(red_wines, quality.lvl==5),cor(residual.sugar,alcohol))
with(subset(red_wines, quality.lvl==6),cor(residual.sugar,alcohol))
with(subset(red_wines, quality.lvl==8),cor(residual.sugar,alcohol))

```

Same as for chlorides, the correlation is quite low. But the trends in the plots for chlorides and residual.sugar look kind of the same. Chlorides and residual sugar might be linked.

###Chlorides and residual sugar by quality

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = chlorides, y = residual.sugar), data = red_wines_quant) +
  geom_jitter(alpha=0.4) +
  geom_smooth(color=I('red')) +
  xlim(quantile(red_wines$chlorides,0.05), 
       quantile(red_wines$chlorides,0.95)) +
  ylim(quantile(red_wines$residual.sugar,0.05), 
       quantile(red_wines$residual.sugar,0.95)) +
  facet_wrap(~quality.lvl, scales = 'free') +
  ggtitle("Chlorides by residual sugar smoother, .95 quantile")

```

The relation between chlorides and residual sugar seems to go more linear as the quality improves. It might make for better balanced wine in the end.

Let see if the ratio of sugar by chlorides is any help:

```{r,echo=FALSE,warning=FALSE}
#adding ratio of residual.sugar/chlorides to red_wines dataframe
red_wines$residual.sugar_chlorides=with(red_wines,residual.sugar/chlorides)

ggplot(aes(x = quality, y = residual.sugar_chlorides), data = red_wines) +
  geom_jitter(alpha = 0.4) +
  ylim(quantile(red_wines$residual.sugar_chlorides,0.05), 
       quantile(red_wines$residual.sugar_chlorides,0.95)) +
  ggtitle("Residual sugar/Chlorides by quality, .95 quantile")

```

No chance here. I was hoping for some clusters of points for each quality level, but they can vary widely.

```{r,warning=FALSE}

with(subset(red_wines, quality.lvl==5),cor(residual.sugar,chlorides))
with(subset(red_wines, quality.lvl==6),cor(residual.sugar,chlorides))
with(subset(red_wines, quality.lvl==8),cor(residual.sugar,chlorides))

```

The correlation for quality 5 and 6 is quite low but the one for 8 is much higher. This might be because we have fewer point on the 8 quality than on other quality level. Furthermore, we have very far off values in both residual.sugar and chlorides, let's calculate the same value on a subset limited between the .05 and .95 quantiles.


```{r,warning=FALSE}
red_wines_quant<-subset(red_wines,
                        quantile(red_wines$residual.sugar,0.05)<=residual.sugar
                        & quantile(red_wines$residual.sugar,0.95)>=residual.sugar
                        & quantile(red_wines$chlorides,0.05)<=chlorides 
                        & quantile(red_wines$chlorides,0.95)>=chlorides)

with(subset(red_wines_quant, quality.lvl==5),
     cor(residual.sugar, chlorides))
with(subset(red_wines_quant, quality.lvl==6),
     cor(residual.sugar, chlorides))
with(subset(red_wines_quant, quality.lvl==8),
     cor(residual.sugar, chlorides))

```

Correlations are much improved with the removing of the extreme quantiles.
I think this relationship between residual sugar and chlorides is worth investigating to see if other variables might come in play.
So let's map chlorides with sugar and add alcohol as a color and use this as a base for other graphs.

###Chlorides, residual sugar and alcohol by quality

```{r,echo=FALSE,warning=FALSE}
#moving to log scale for better view and adding smoother

sugar.chlorides<-ggplot(aes(x = residual.sugar, y = chlorides), 
                        data = red_wines_quant) +
  facet_wrap(~quality.lvl, scales = 'free') +
  scale_y_log10() +
  scale_x_log10()
  
sugar.chlorides+
  geom_jitter(aes(color=alcohol)) +
  scale_colour_gradient2(low="red", high="blue", midpoint=11) +
  ggtitle("Chlorides by residual sugar and alcohol, log axes")

```

We can see an upward trend and that low alcohol are most of the time in the upper left whereas higher alcohol wines are in the bottom right.

```{r,warning=FALSE}
with(red_wines_quant, cor(residual.sugar_chlorides, alcohol))
with(subset(red_wines_quant, quality.lvl==5),
     cor(residual.sugar_chlorides, alcohol))
with(subset(red_wines_quant, quality.lvl==6),
     cor(residual.sugar_chlorides, alcohol))
with(subset(red_wines_quant, quality.lvl==8),
     cor(residual.sugar_chlorides, alcohol))
```

The correlation is medium low and does not seem to vary greatly between qualities.

###Acidity variables and quality
Let's go back to our acidity variables:

```{r,echo=FALSE,warning=FALSE}
ggplot_base_acidity<-ggplot(aes(x = quality.lvl), data = red_wines)

#fixed acidity
ggplot_base_acidity +
  geom_boxplot(aes(y = fixed.acidity)) +
  ggtitle("Fixed acidity by quality")

#volatile acidity
ggplot_base_acidity +
  geom_boxplot(aes(y = volatile.acidity)) +
  ggtitle("Volatile acidity by quality")

#citric acidity
ggplot_base_acidity +
  geom_boxplot(aes(y = citric.acid)) +
  ggtitle("Citric acid by quality")
```

We can see that fixed acidity is not really relevant to quality, but volatile.acidity and citric acid are. The higher the citric acid level, the better the quality. The relation between volatile acidity and quality is the other way around. Futhermore, looking at the median we can see some group of qualities: 3 & 4, 5 & 6 and 7 & 8.

```{r,warning=FALSE}
with(red_wines, cor(quality, fixed.acidity))
with(red_wines, cor(quality, volatile.acidity))
with(red_wines, cor(quality, citric.acid))
```

Correlation confirms that fixed acidity correlates only weakly with the quality.
On the other hand, volatile acidity and citric acid are much stronger, especially volatile acidity. This kind of acidity correlate negatively with quality meaning that as the quality improves, the volatile acidity decreases.


So we now have alcohol, citric acid and volatile acidity correlating quite strongly with quality.
I will get back to the previous facetted chart of residual sugar and chlorides and add citric acid/volatile acidity to see if I can get some more information.

###Chlorides, residual sugar and acidities by quality

```{r,echo=FALSE,warning=FALSE}
#volatile acidity
sugar.chlorides+
  geom_jitter(aes(color=volatile.acidity)) +
  scale_colour_gradient2(low="red", high="blue", midpoint=0.52) +
  ggtitle('Chlorides, residual sugar & volatile acidity facetted by quality')

#citric acid
sugar.chlorides+
  geom_jitter(aes(color=citric.acid)) +
  scale_colour_gradient2(low="red", high="blue", midpoint=0.25) +
  ggtitle('Chlorides, residual sugar & citric acid facetted by quality')

```

Nothing for volatile acidity. But it seems that we get more citric acid the more residual sugar and chlorides in the wine.

```{r,warning=FALSE}
with(red_wines_quant, cor(residual.sugar_chlorides, volatile.acidity))
with(red_wines_quant, cor(residual.sugar_chlorides, citric.acid))
with(subset(red_wines_quant, quality.lvl==5),
     cor(residual.sugar_chlorides, citric.acid))
with(subset(red_wines_quant, quality.lvl==6),
     cor(residual.sugar_chlorides, citric.acid))
with(subset(red_wines_quant, quality.lvl==8),
     cor(residual.sugar_chlorides, citric.acid))
```

Correlations show that there is in fact nothing in there.

###Chlorides, residual sugar and density by quality
Density is related to alcohol so let see if we can find something here.

```{r,echo=FALSE,warning=FALSE}
sugar.chlorides+
  geom_jitter(aes(color=density)) +
  scale_colour_gradient2(low="red", high="blue", midpoint=0.9975) +
  ggtitle("Chlorides by residual sugar and density, log axes")

```

We get the same kind of graph than for alcohol with denser wines on the right of the graphs whereas less dense wines are in the left region. 

```{r,warning=FALSE}
with(red_wines_quant, cor(residual.sugar_chlorides, density))
with(red_wines_quant, cor(residual.sugar_chlorides, alcohol))
with(subset(red_wines_quant, quality.lvl==5),
     cor(residual.sugar_chlorides, density))
with(subset(red_wines_quant, quality.lvl==5),
     cor(residual.sugar_chlorides, alcohol))

```

Density relates less to the residual.sugar_chlorides ratio than alcohol.

###Chlorides, residual sugar and sulphates by quality

```{r,echo=FALSE,warning=FALSE}
#adding a limit on the sulphates as there seems to be outlier far off
sugar.chlorides+
  geom_jitter(aes(color=sulphates), 
              data = subset(red_wines, 
                            sulphates<quantile(red_wines$sulphates,0.95))) +
  scale_colour_gradient2(low="red", high="blue", midpoint=0.6) +
  ggtitle("Chlorides by residual sugar and sulphates, log axes")

#seems to have more sulphates for higher quality wines
ggplot(aes(x = quality.lvl, y = sulphates), 
       data = subset(red_wines, 
                     sulphates<quantile(red_wines$sulphates,0.95))) +
  geom_boxplot() +
  ggtitle("Sulphates by quality, with sulphates<0.75)")

```

```{r,warning=FALSE}
with(red_wines_quant, cor(residual.sugar_chlorides, sulphates))
with(subset(red_wines_quant, quality.lvl==5),
     cor(residual.sugar_chlorides, sulphates))
with(subset(red_wines_quant, quality.lvl==6),
     cor(residual.sugar_chlorides, sulphates))
with(subset(red_wines_quant, quality.lvl==8),
     cor(residual.sugar_chlorides, sulphates))
```

The higher the quality, the more sulphates there is. Otherwise, no relation with chlorides and residual sugar. Correlations by quality are low and increases for quality 8 but this quality has only a few data points.

Let's now look at sulfurs.

###Chlorides, residual sugar and sulfurs by quality

```{r,echo=FALSE,warning=FALSE}
summary(red_wines$free.sulfur.dioxide)
#subset to filter outlier
      
sugar.chlorides+
  geom_jitter(aes(color=free.sulfur.dioxide), 
              data = subset(red_wines, 
                     free.sulfur.dioxide<quantile(red_wines$free.sulfur.dioxide,
                                                  0.95))) +
  scale_colour_gradient2(low="red", high="blue", midpoint=14) +
  ggtitle('chlorides by residual sugar & 
          free sulfur dioxide facetted by quality')

summary(red_wines$total.sulfur.dioxide)
#subset to filter outlier
sugar.chlorides+
  geom_jitter(aes(color=total.sulfur.dioxide), 
              data = subset(red_wines, 
                     total.sulfur.dioxide < 
                       quantile(red_wines$total.sulfur.dioxide, 0.95))) +
  scale_colour_gradient2(low="red", high="blue", midpoint=38) +
  ggtitle('chlorides by residual sugar & total sulfur dioxide facetted by quality')
```

```{r,warning=FALSE}
with(red_wines_quant, cor(residual.sugar_chlorides, free.sulfur.dioxide))
with(red_wines_quant, cor(residual.sugar_chlorides, total.sulfur.dioxide))
```

Sulfur variables do not correlate with chlorides and residual sugar.

#Final Plots and Summary
First of all, there are outliers in the dataset. I tried to remove them when their value seemed to really be out of range. But it might also underlines the diversity of wines. As I am not sure, the following graphs are based on the whole dataset.
All in all, there is not some really strong correlations between quality and the other variables presented here. But I could understood some relationships regarding tastes:

##Vinegar taste is not really looked for

```{r,echo=FALSE,warning=FALSE}

ggplot(aes(x = volatile.acidity, fill=quality.lvl), data = red_wines) +
  geom_density(aes(y=..scaled..)) +
  scale_fill_brewer(type = 'seq', palette = 3) +
  labs(x = 'Volatile acidity (g / dm^3)', 
       y = 'Scaled density',  
       fill='Quality') +
  ggtitle("Volatile acidity by quality")

```  
  
The density were scaled so that the low counts of the extreme quality do affect the overall distribution.
As the quality of the wines get higher, the volatile acidity is getting lower. It is not always true. This is kind of what we can expect. When you buy cheap wine (which is not always, but most of the time of lower quality) it has most of the tume a pugent smell and taste, like vinegar.
  
##But citric acid is the other way around

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = quality.lvl, y = citric.acid), data = red_wines) +
  geom_boxplot(outlier.colour = 'red') +  
  labs(x = 'Quality level', 
       y = 'Citric acid (g / dm^3)') +
  ggtitle("Citric Acid by quality (red outliers)")
```

This kind of acid add freshness and flavor. It is confirmed by the boxplot: lower quality wines have less citric acid in general than their higher contourparts. We can group quality into 3 supra group:

1. Quality 3 & 4: low citric acid median at `r with(subset(red_wines, red_wines$quality.lvl==3 | red_wines$quality.lvl==4), median(citric.acid))`
2. Quality 5 & 6: median around `r with(subset(red_wines, red_wines$quality.lvl==5 | red_wines$quality.lvl==6), median(citric.acid))`
3. Quality 7 & 8: median around `r with(subset(red_wines, red_wines$quality.lvl==7 | red_wines$quality.lvl==8), median(citric.acid))`

##Get high alcohol level and  high sulphates for better wines

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = sulphates, y = alcohol, color = quality.lvl), 
       data = red_wines) +
  stat_smooth(level=0.75) +
  coord_cartesian() +
  xlim(0.4,0.75) +
  ylim(6,15) +
  scale_color_brewer(type = 'seq', palette = 3) +
  labs(x = 'Sulphates (g / dm3)', 
       y = 'Alcohol (%/volume)', 
       color = 'Quality') +
  ggtitle("Alcohol by sulphates and quality (.75 confidence interval)")
```

This graph shows smoothers of alcohol by sulphates for all quality. The low quality (3) wines vary a lot but the overal trend is that they are high in alcohol with low sulphates and the alcohol level decreases as the sulphates level increases. Qualities 4 to 7 are flat overall and the only differences come from the alcohol level.
Quality 8 wines are peculiar in that they are grouped in the top right corner. The relationship is not flat like the other quality (except 3) but those  wines have more alcohol and more sulphates. 


#Reflection
I started the analysis by looking at the data overal. Some data points seemed really far out and during the analysis, I removed some of them. But I decided to keep them in the end as wine tastes can vary greatly.  
Looking at the different variables, I decidied to investigate what made a wine good.
I  started by looking at the main things that could affect taste: alcohol, chlorides, residual.sugar and  pH. My first finding was really about alcohol and quality.  
I then moved to multivariables analyses by combining alcohol with chlorides and residual.sugar. The overall trends by quality looked quite similar so I tried to plot chlorides with residual.sugar. But the results was not what I was expecting. Thinking that another factor might be related with those two, I went through the the acidity variables and the quality to select volatile.acidity and citric.acidity and plot them with chlorides and residual.sugar. The plots seemed to give something for citric.acid byt I was overreading it, the correlation is low.
I did the same exercice for the rest of the variables.  
I get all in all several variables correlating with quality:
* Alcohol
* Volatile acidity (negative correlation)
* Citric acid
* Sulphates
I was looking for some more adavanced relations between the variables but was not able to find them. I am especially surprised that residual.sugar and chlorides does not affect quality more.  
The limits of the dataset are really the lack of points on the lower and higher quality wines. Futhermore, the source of the quality is unknown, is it from a professionnal? A store? And we have to keep in mind that taste is a really cultural thing (see the [indian food analysis](http://www.washingtonpost.com/blogs/wonkblog/wp/2015/03/03/
a-scientific-explanation-of-what-makes-indian-food-so-delicious/)) and a good wine for someone might not be for another.
Last but not least, we do not have the age of the wine. In France, one of the first thing we check for a wine is its age because common knowledge dictates that an older wine is better than a younger one (for "normal" year). I would have love to validate or discard this coutume.
Another thing that I would like is to get the name of the wines so that I can get their prices and see if how they relate to quality.